The hydra and the rooster

### 2020-06-30 latex/chapter-ks.tex
standradizing order of arguments of function canon and derivates

Our definition may look more complex than the mathematical one, but
uses plain structural recursion over the type \coqsimple{T1}. Thus, tactics like
\coqsimple{cbn}, \coqsimple{simpl}, \coqsimple{compute}, etc., are applicable. For simplicity's sake, 
we used an auxiliary function \texttt{canonS} of type \texttt{nat -> T1 -> T1} such that
\texttt{canonS i $\alpha$} is equal to $\canonseq{\alpha}{i+1}$.

@@

\index{Functions!canonS canonS (helper for canon)}
\index{Functions!canon @ canon (canonical sequence)}
\begin{Coqsrc}
Fixpoint canonS (i:nat) alpha :=
  match alpha with
      zero => zero
    | ocons zero 0 zero => zero
    | ocons zero (S k) zero => FS k
    | ocons gamma 0 zero =>
      match pred gamma with
          Some gamma' => ocons gamma' i zero
        | None => ocons (canonS i gamma) 0 zero
      end
    | ocons gamma (S n) zero =>
       match pred gamma with
           Some gamma' => ocons gamma n (ocons gamma' i zero)
         | None => ocons gamma n (ocons (canonS i gamma) 0 zero)
       end
    | ocons alpha n beta => ocons alpha n (canonS i beta)
  end.
\end{Coqsrc}

@@

\begin{Coqsrc}
 Definition canon i alpha := 
   match i with 0 => zero | S j => canonS j alpha end. 
\end{Coqsrc}

@@

\begin{Coqsrc}
Compute (canon 3 (omega ^ omega)).
\end{Coqsrc}

@@

\begin{Coqsrc}
Example canon3 :  canon 3 (omega ^ omega) = omega ^ 3.
Proof. reflexivity. Qed.
\end{Coqsrc}

\begin{Coqsrc}
Compute pp (canon 42 (omega ^ omega * 3)).  
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma canonS_succ i alpha :
  nf alpha ->  canonS i (succ alpha) = alpha.
Proof.
 induction alpha.
 (* ... *)
\end{Coqsrc}

@@

\label{lemma:canonS_LT}
\begin{Coqsrc}
Lemma canonS_LT i alpha :
  nf alpha -> alpha <> zero -> canonS i alpha <  alpha.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma canonS_limit_strong (lambda : T1) : 
     nf lambda ->
     is_limit lambda  ->
     forall beta, beta < lambda ->
                  {i:nat | beta < canonS i lambda}.

@@

\begin{Coqsrc}
Lemma canonS_limit_lub (lambda : T1) :
  nf lambda -> is_limit lambda  ->
  strict_lub (fun i => canonS i lambda) lambda.
\end{Coqsrc}

@@

\begin{Coqsrc}
(Definition transition_S i : relation T1 :=
  fun alpha beta =>  alpha <> zero /\ beta = canonS i alpha.

@@

\begin{Coqsrc}
Lemma canonS_iota i alpha :
    nf alpha -> alpha <> 0 ->
    iota alpha -1-> iota (canonS i alpha).
\end{Coqsrc}

@@

Inductive standard_pathR(j:nat)( beta:T1):  nat -> T1 -> Prop :=
  std_1 : forall i alpha, 
       beta = canon i alpha -> j = S i ->
       standard_pathR j beta i  alpha
| std_S : forall i alpha, 
      standard_pathR j beta (S i) (canon i alpha)  ->
      standard_pathR j beta i alpha.

@@

\begin{Coqsrc}
Definition const_pathS i :=
    clos_trans_1n T1 (fun alpha beta => beta = canonS i alpha).

@@

  \begin{Coqsrc}
Theorem Theorem_2_4 (lambda : T1) :
   nf lambda ->
   is_limit lambda  ->
   forall i j, (i < j)%nat ->
               const_pathS 0 (canonS j lambda)
                             (canonS i lambda). 
  \end{Coqsrc}

@@

\begin{Coqsrc}
Fixpoint standard_gnaw (i:nat)(alpha : T1)(l:nat):  T1  :=
  match l with
  | 0 => alpha
  | S m => standard_gnaw (S i) (canon i alpha) m
  end.
\end{Coqsrc}

>>>>

Our definition may look more complex than the mathematical one, but
uses plain structural recursion over the type \coqsimple{T1}. Thus, tactics like
\coqsimple{cbn}, \coqsimple{simpl}, \coqsimple{compute}, etc., are applicable. For simplicity's sake, 
we used an auxiliary function \texttt{canonS} of type \texttt{T1 -> nat  -> T1} such that
\texttt{canonS  $\alpha$ $i$} is equal to $\canonseq{\alpha}{i+1}$.

@@

\index{Functions!canonS canonS (helper for canon)}
\index{Functions!canon @ canon (canonical sequence)}
\begin{Coqsrc}
Fixpoint canonS alpha (i:nat) :=
  match alpha with
      zero => zero
    | ocons zero 0 zero => zero
    | ocons zero (S k) zero => FS k
    | ocons gamma 0 zero =>
      match pred gamma with
          Some gamma' => ocons gamma' i zero
        | None => ocons (canonS gamma i) 0 zero
      end
    | ocons gamma (S n) zero =>
       match pred gamma with
           Some gamma' => ocons gamma n (ocons gamma' i zero)
         | None => ocons gamma n (ocons (canonS gamma i) 0 zero)
       end
    | ocons alpha n beta => ocons alpha n (canonS beta i)
  end.
\end{Coqsrc}

@@

\begin{Coqsrc}
 Definition canon alpha i := 
   match i with 0 => zero | S j => canonS alpha j end. 
\end{Coqsrc}

@@

\begin{Coqsrc}
Compute (canon (omega ^ omega) 3).
\end{Coqsrc}

@@

\begin{Coqsrc}
Example canon3 :  canon (omega ^ omega) 3 = omega ^ 3.
Proof. reflexivity. Qed.
\end{Coqsrc}

\begin{Coqsrc}
Compute pp (canon (omega ^ omega * 3) 42).  
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma canonS_succ i alpha :
  nf alpha ->  canonS (succ alpha) i = alpha.
Proof.
 induction alpha.
 (* ... *)
\end{Coqsrc}

@@

\label{lemma:canonS_LT}
\begin{Coqsrc}
Lemma canonS_LT i alpha :
  nf alpha -> alpha <> zero -> canonS alpha i <  alpha.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma canonS_limit_strong (lambda : T1) : 
     nf lambda ->
     is_limit lambda  ->
     forall beta, beta < lambda ->
                  {i:nat | beta < canonS lambda i}.

@@

\begin{Coqsrc}
Lemma canonS_limit_lub (lambda : T1) :
  nf lambda -> is_limit lambda  ->
  strict_lub (fun i => canonS lambda i) lambda.
\end{Coqsrc}

@@

\begin{Coqsrc}
(Definition transition_S i : relation T1 :=
  fun alpha beta =>  alpha <> zero /\ beta = canonS alpha i.

@@

\begin{Coqsrc}
Lemma canonS_iota i alpha :
    nf alpha -> alpha <> 0 ->
    iota alpha -1-> iota (canonS alpha i).
\end{Coqsrc}

@@

Inductive standard_pathR(j:nat)( beta:T1):  nat -> T1 -> Prop :=
  std_1 : forall i alpha, 
       beta = canon alpha i -> j = S i ->
       standard_pathR j beta i  alpha
| std_S : forall i alpha, 
      standard_pathR j beta (S i) (canon alpha i)  ->
      standard_pathR j beta i alpha.

@@

\begin{Coqsrc}
Definition const_pathS i :=
    clos_trans_1n T1 (fun alpha beta => beta = canonS alpha i).

@@

  \begin{Coqsrc}
Theorem Theorem_2_4 (lambda : T1) :
   nf lambda ->
   is_limit lambda  ->
   forall i j, (i < j)%nat ->
               const_pathS 0 (canonS lambda j)
                             (canonS lambda i). 
  \end{Coqsrc}

@@

\begin{Coqsrc}
Fixpoint standard_gnaw (i:nat)(alpha : T1)(l:nat):  T1  :=
  match l with
  | 0 => alpha
  | S m => standard_gnaw (S i) (canon alpha i) m
  end.
\end{Coqsrc}

### 2020-06-30 latex/chapter-large-sets.tex
standradizing order of arguments of function canon and derivates

\begin{Coqsrc}
canonS_LT:
  forall (i : nat) (alpha : T1),
  nf alpha -> alpha <> zero -> canonS i alpha < alpha.

@@

\begin{Coqsrc}
Section lim.
  Variables (lambda : T1)
            (Hnf : nf lambda)
            (Hlim : is_limit lambda)
            (f : nat -> nat -> nat)
            (H : forall k, L_spec (canonS k lambda) (f (S k))).

@@

Equations  L_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
  L_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Is_Limit alpha) :=
          { | left _ =>  L_ (Canon i alpha)  (S i) ;
            | right notlimit =>  L_ (Pred alpha) (S i)}}.

@@

Lemma L_lim_eqn alpha i : Is_Limit alpha -> L_ alpha i =
                                        L_ (Canon i alpha) (S i).

@@

\begin{Coqsrc}
Equations H_ (alpha: E0) (i:nat) :  nat  by wf  alpha lt :=
  H_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Is_Limit alpha) :=
          { | left _ =>  H_ (Canon (S i) alpha)  i ;
            | right notlimit =>  H_ (Pred alpha) (S i)}}.

@@

Lemma H_lim_eqn alpha i : Is_Limit alpha ->
                      H_ alpha i =  H_ (Canon (S i) alpha) i.

@@

\begin{Coqsrc}
Lemma H_omega : forall k, H_ Omega k = S (2 * k)%nat.
Proof with auto with E0.
  intro k; rewrite H_lim_eqn ...
  - replace (Canon (S k) omega) with (Fin (S k)).
    + rewrite H_Fin; lia.
    +  now autorewrite with E0_rw.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma H_omega_double k : H_ (omega * 2)%e0 k =  (4 * k + 3)%nat.
Proof.
 rewrite H_lim_eqn; simpl Canon.
 - ochange  (CanonS  k (omega * 2)%e0)  (omega + (S k))%e0.
  + rewrite H_Plus_Fin, H_omega;  lia.
  -  now compute.
Qed.

@@

Lemma H_omega_sqr : forall k,
    H_ (Phi0  2)%e0 k = (exp2 (S k ) *  S k - 1)%nat.
Proof.
  intro k; 
   rewrite H_lim_eqn; auto with E0.
  - ochange (Canon (S k) (Phi0 2)) (omega * (S k))%e0.
    +  rewrite H_omega_i; simpl (exp2 (S k)).
       *  rewrite Nat.add_pred_r.
          -- lia. 
          --   generalize (exp2_not_zero k);  lia.
    + cbn; f_equal; lia.
Qed.
\end{Coqsrc}

@@

\index{Coq!Plug-ins!Equations}
\begin{Coqsrc}
Fail Equations F_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
  F_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Is_Limit alpha) :=
          { | left _ =>  F_ (Canon i alpha)  i ;
            | right notlimit =>  iterate (F_ (Pred alpha))  (S i) i}}.
\end{Coqsrc}

@@

Equations  F_star (c: E0 * nat) (i:nat) :  nat by wf  c call_lt :=
    F_star (alpha, 0) i := i;
    F_star (alpha, 1) i
      with E0_eq_dec alpha Zero :=
           { | left _ => S i ;
             | right nonzero
                 with Utils.dec (Is_Limit alpha) :=
                 { | left _ => F_star (Canon i alpha,1) i ;
                   | right notlimit =>
                     F_star (Pred alpha, S i)  i}};
    F_star (alpha,(S (S n))) i :=
               F_star (alpha, 1) (F_star (alpha, (S n)) i).

@@

Lemma F_lim_eqn : forall alpha i,  Is_Limit alpha ->
                               F_ alpha i = F_ (Canon i alpha) i.

>>>>

\begin{Coqsrc}
canonS_LT:
  forall (i : nat) (alpha : T1),
  nf alpha -> alpha <> zero -> canonS alpha i < alpha.

@@

\begin{Coqsrc}
Section lim.
  Variables (lambda : T1)
            (Hnf : nf lambda)
            (Hlim : is_limit lambda)
            (f : nat -> nat -> nat)
            (H : forall k, L_spec (canonS lambda k) (f (S k))).

@@

Equations  L_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
  L_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Is_Limit alpha) :=
          { | left _ =>  L_ (Canon alpha i)  (S i) ;
            | right notlimit =>  L_ (Pred alpha) (S i)}}.

@@

Lemma L_lim_eqn alpha i : Is_Limit alpha -> L_ alpha i =
                                        L_ (Canon alpha i) (S i).

@@

\begin{Coqsrc}
Equations H_ (alpha: E0) (i:nat) :  nat  by wf  alpha lt :=
  H_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Is_Limit alpha) :=
          { | left _ =>  H_ (Canon alpha (S i))  i ;
            | right notlimit =>  H_ (Pred alpha) (S i)}}.

@@

Lemma H_lim_eqn alpha i : Is_Limit alpha ->
                      H_ alpha i =  H_ (Canon alpha (S i)) i.

@@

\begin{Coqsrc}
Lemma H_omega : forall k, H_ Omega k = S (2 * k)%nat.
Proof with auto with E0.
  intro k; rewrite H_lim_eqn ...
  - replace (Canon omega (S k)) with (Fin (S k)).
    + rewrite H_Fin; lia.
    +  now autorewrite with E0_rw.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma H_omega_double k : H_ (omega * 2)%e0 k =  (4 * k + 3)%nat.
Proof.
 rewrite H_lim_eqn; simpl Canon.
 - ochange  (CanonS  (omega * 2)%e0 k)  (omega + (S k))%e0.
  + rewrite H_Plus_Fin, H_omega;  lia.
  -  now compute.
Qed.

@@

Lemma H_omega_sqr : forall k,
    H_ (Phi0  2)%e0 k = (exp2 (S k ) *  S k - 1)%nat.
Proof.
  intro k; 
   rewrite H_lim_eqn; auto with E0.
  - ochange (Canon (Phi0 2) (S k)) (omega * (S k))%e0.
    +  rewrite H_omega_i; simpl (exp2 (S k)).
       *  rewrite Nat.add_pred_r.
          -- lia. 
          --   generalize (exp2_not_zero k);  lia.
    + cbn; f_equal; lia.
Qed.
\end{Coqsrc}

@@

\index{Coq!Plug-ins!Equations}
\begin{Coqsrc}
Fail Equations F_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
  F_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Is_Limit alpha) :=
          { | left _ =>  F_ (Canon alpha i)  i ;
            | right notlimit =>  iterate (F_ (Pred alpha))  (S i) i}}.
\end{Coqsrc}

@@

Equations  F_star (c: E0 * nat) (i:nat) :  nat by wf  c call_lt :=
    F_star (alpha, 0) i := i;
    F_star (alpha, 1) i
      with E0_eq_dec alpha Zero :=
           { | left _ => S i ;
             | right nonzero
                 with Utils.dec (Is_Limit alpha) :=
                 { | left _ => F_star (Canon alpha i,1) i ;
                   | right notlimit =>
                     F_star (Pred alpha, S i)  i}};
    F_star (alpha,(S (S n))) i :=
               F_star (alpha, 1) (F_star (alpha, (S n)) i).

@@

Lemma F_lim_eqn : forall alpha i,  Is_Limit alpha ->
                               F_ alpha i = F_ (Canon alpha i) i.
